---
output:
  md_document:
    variant: markdown_github
---



# User friendly extensions to Project Gutenberg

> I remembered the glowworms and millipeds.  
> I remembered the spectre.  
> I remembered the dressmaker's smile!! 



## Intro

Working with raw text files from Project Gutenberg. Some simple, parallelized functions for text extraction, sentence tokenization, and simple search. Extensions assume corpus has been downloaded, lives locally.

See [this repository](https://github.com/pgcorpus/gutenberg) for a quick/easy download of full Project Gutenberg corpus.  No need to process the raw files.  Output includes a nice meta file summarizing text included in download.


```{r include=FALSE}
path_to_raw_pg_files <- '/home/jtimm/gutenberg/data/raw/'
path_to_out_folder <- '/home/jtimm/gutenberg/data/sentences/'
```




## Extraction & sentence tokenization


### Extraction function

Taken from [this R package](https://github.com/patperry/r-corpus/blob/master/R/gutenberg.R), and tweaked ever so slightly.

```{r}
pg_extract_text <- function(lines){
    # number the lines
    i <- seq_along(lines)

    # find the empty lines
    empty <- (lines == "")

    # find the end of the Project Gutenberg header
    start_pat <- paste("^[*][*][*].*PROJECT GUTENBERG.*[*][*][*]",
                       "END.*SMALL PRINT", sep = "|")
    start_match <- grep(start_pat, lines)
    if (length(start_match) == 0) {
        start_match <- 0
    }
    start <- start_match[[1]] + 1

    # look for the encoding, and convert to UTF-8
    enc_pat <- "Character set encoding:[[:space:]]*(.*)[[:space:]]*$"
    enc_match <- grep(enc_pat, lines[seq_len(start - 1)])
    if (length(enc_match) > 0) {
        enc <- sub(enc_pat, "\\1", lines[enc_match[[1]]])
        if (!enc %in% c("ASCII", "UTF-8")) {

          tryCatch({
            lines <- iconv(lines, enc, "UTF-8")},
            error = function(e) {lines = "REMOVE"})
        }
    }

    start <- min(which(start <= i & !empty))

    # find the start of the Project Gutenberg footer
    end_pat <- paste("^End of .*Project Gutenberg.*",
                     "\\\\*\\\\*\\\\*.*END OF.*PROJECT GUTENBERG", sep = "|")

    end_match <- grep(end_pat, lines)
    if (length(end_match) == 0) {
        end_match <- length(lines) + 1
    }
    end <- end_match[[1]] - 1

    # skip the empty lines at the end
    end <- max(which(i <= end & !empty))

    # skip the production notes at the start of the text
    note_start_pat <- paste("produced by",
                            "prepared by",
                            "transcribed from",
                            "project gutenberg",
                            "^[*][*][*]",
                            "^note: ",
                            "^special thanks",
                            "^this is a retranscription",
                            sep = "|")
    note_start <- grep(note_start_pat, lines, ignore.case = TRUE)
    note_start <- note_start[start <= note_start & note_start <= end]

    ## error happens in here -- 
    while (length(note_start) && note_start[[1]] == start) {
        # the note ends at the first empty line
        note_end <- min(which(note_start[[1]] <= i & empty))

        start <- min(which(note_end + 1 <= i & !empty))
        note_start <- note_start[start <= note_start]
    }

    # concatenate the content lines
    if(!is.finite(start)){'REMOVE'} else{
      paste(lines[start:end], collapse = "\n")
     }
}
```




### Sentence tokenization & output function

```{r}
pg_tokenize_sentences <- function(x){
  
  lapply(x, function(y){
    id <- gsub('^.*/', '',  y)
    id <- gsub('_.*$', '', id)
    lines <- readLines(y, 
                       encoding = "UTF-8", 
                       warn = FALSE) 
    
    lines0 <- pg_extract_text(lines)
    
    if(lines[1] == 'REMOVE'){} else{
      
      tryCatch({
        x0 <- data.frame(doc_id = id,
                         text = lines0) |>
          text2df::tif2sentence() |> 
          data.table::setDT()
        
        x0[, text := trimws(gsub('\n', ' ', text))]
        #setwd(out_dir)
        saveRDS(x0, paste0(id, '.rds'))
        },
        
        error = function(e) {})
      }
  })
}
```



### Batches, parallel

```{r eval=FALSE}
pafs <- list.files(path = path_to_raw_pg_files,
                   full.names = T)

batches <- split(pafs, ceiling(seq_along(pafs)/50))

setwd(path_to_out_folder)

clust <- parallel::makeCluster(7)
parallel::clusterExport(cl = clust,
                        varlist = c('pg_tokenize_sentences',
                                    'pg_extract_text',
                                    'batches'),
                        envir = environment())

pbapply::pblapply(X = batches, 
                  FUN = pg_tokenize_sentences, 
                  cl = clust)

parallel::stopCluster(clust)
```



```{r eval=FALSE, include=FALSE}
setwd(path_to_meta)
pg_meta <- read.csv('metadata.csv')

pg_meta <- pg_meta |>
  mutate(subjects = gsub('\\{|\\}', '', subjects)) |>
  mutate(subjects = gsub('"', "'", subjects)) |>
  tidyr::separate_rows(subjects, sep = "',") |>
  mutate(subjects = gsub("^'|'$|^ '", "", subjects)) |>
  mutate(subjects = trimws(subjects)) |> 
  group_by(across(c(-subjects))) |>
  summarize(subjects = paste0(subjects, collapse = ' | ')) |>
  ungroup()

pg_subjects <- pg_meta |>
  select(id, authoryearofbirth, language, subjects) |>
  tidyr::separate_rows(subjects, sep = " \\| ")

setwd('/home/jtimm/pCloudDrive/GitHub/packages/project-gutenberg')
usethis::use_data(pg_meta, overwrite=TRUE)
usethis::use_data(pg_subjects, overwrite=TRUE)
```





## Package

```{r eval=FALSE}
devtools::install_github("jaytimm/gbr")
```


### Metadata

```{r}
subs <- gbr::gb_subset(x = 10,
                       min_birth_year = 1850,
                       max_birth_year = 1900,
                       subj = 'Science fiction',
                       lang = 'en',
                       filepath = path_to_out_folder) 
```



### Build corpus

```{r message=FALSE, warning=FALSE}
subs |> 
  gbr::gb_build_corpus() |>
  dplyr::sample_n(5) |> 
  knitr::kable()
```




### Simple regex search

```{r message=FALSE, warning=FALSE}
gbr::gb_subset(x = 5000) |>
  gbr::gb_search(pattern = '^I remembered', cores = 12) |>
  gbr::gb_examples(n = 1) |>
  knitr::kable()
```


```{r eval=FALSE, include=FALSE}
# Oh, where have you been, my blue-eyed son?
# Oh, where have you been, my darling young one?
# I’ve stumbled on the side of twelve misty mountains
# I’ve walked and I’ve crawled on six crooked highways
# I’ve stepped in the middle of seven sad forests
# I’ve been out in front of a dozen dead oceans
# I’ve been ten thousand miles in the mouth of a graveyard
# And it’s a hard, and it’s a hard, it’s a hard, and it’s a hard
# And it’s a hard rain’s a-gonna fall
# 
# Oh, what did you see, my blue-eyed son?
# Oh, what did you see, my darling young one?
# I saw a newborn baby with wild wolves all around it
# I saw a highway of diamonds with nobody on it
# I saw a black branch with blood that kept drippin’
# I saw a room full of men with their hammers a-bleedin’
# I saw a white ladder all covered with water
# I saw ten thousand talkers whose tongues were all broken
# I saw guns and sharp swords in the hands of young children
# And it’s a hard, and it’s a hard, it’s a hard, it’s a hard
# And it’s a hard rain’s a-gonna fall
# 
# And what did you hear, my blue-eyed son?
# And what did you hear, my darling young one?
# I heard the sound of a thunder, it roared out a warnin’
# Heard the roar of a wave that could drown the whole world
# Heard one hundred drummers whose hands were a-blazin’
# Heard ten thousand whisperin’ and nobody listenin’
# Heard one person starve, I heard many people laughin’
# Heard the song of a poet who died in the gutter
# Heard the sound of a clown who cried in the alley
# And it’s a hard, and it’s a hard, it’s a hard, it’s a hard
# And it’s a hard rain’s a-gonna fall
# 
# Oh, who did you meet, my blue-eyed son?
# Who did you meet, my darling young one?
# I met a young child beside a dead pony
# I met a white man who walked a black dog
# I met a young woman whose body was burning
# I met a young girl, she gave me a rainbow
# I met one man who was wounded in love
# I met another man who was wounded with hatred
# And it’s a hard, it’s a hard, it’s a hard, it’s a hard
# It’s a hard rain’s a-gonna fall
# 
# Oh, what’ll you do now, my blue-eyed son?
# Oh, what’ll you do now, my darling young one?
# I’m a-goin’ back out ’fore the rain starts a-fallin’
# I’ll walk to the depths of the deepest black forest
# Where the people are many and their hands are all empty
# Where the pellets of poison are flooding their waters
# Where the home in the valley meets the damp dirty prison
# Where the executioner’s face is always well hidden
# Where hunger is ugly, where souls are forgotten
# Where black is the color, where none is the number
# And I’ll tell it and think it and speak it and breathe it
# And reflect it from the mountain so all souls can see it
# Then I’ll stand on the ocean until I start sinkin’
# But I’ll know my song well before I start singin’
# And it’s a hard, it’s a hard, it’s a hard, it’s a hard
# It’s a hard rain’s a-gonna fall										
```






